---
title: "Timeseries"
author: "Jyosmitha"
date: "8/13/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
#Importing the Library
library(forecast)
library(tseries)
library(xts)
library(Metrics)
library(ggplot2)
#reading the dataset
data<- forecast::gas
```

```{r}
#Structure of the data
str(data)
#Summary of entire Series
summary(data)
#Checking for null values
anyNA(data)
#Checking the cycle/periodicity of the data
cycle(data)
#frequency of time series
frequency(data)
#View(data)
```

```{r}
#head and tail
head(data)
tail(data)
#start and end of time series
start(data)
end(data)
#outliers
tsoutliers(data)
```

```{r}
#plotting time series
plot(data,main="TimeSeries plot for Australian Gas Production Data",
     col="blue",
     xlab="Year",
     ylab="Gas Production")
#regression line
abline(reg=lm(data~time(data)),col='red')
```

```{r}
#aggregation
data_yearly=aggregate(data,nfrequency = 1)
data_Quarterly=aggregate(data,nfrequency = 4)

#Seasonal Plot
seasonplot(data,year.labels = TRUE,year.labels.left = TRUE,pch=19,col=1:40,
           main='Australian Monthly Gas Production-Seasonal Plot')
#Monthly Plot
monthplot(data,pch=19,xlab='Month',ylab='Gas Production',
          main='Australian Monthly Gas Production-Seasonal Plot',col='red')
#Box plot
boxplot(data~cycle(data),xlab = 'Month'
        , ylab = "Production",main="Boxplot for Australian Gas Production")

ggseasonplot(data,polar = TRUE)
gglagplot(data,main="Lagplot")
```


```{r}
periodicity(data)
```


```{r}
#Decomposing the data into trend,seasonality and randomness
decomAdd=decompose(data,type = "additive")
plot(decomAdd)
decomMul=decompose(data,type="multiplicative")
plot(decomMul)

```

```{r}
#plotting each of the above
plot(decomAdd$trend,xlab="Year",main="Trend Chart after Decomposition")
plot(decomAdd$seasonal,xlab="Year",main="Seasonality Chart after Decomposition")
plot(decomAdd$random,xlab="Year",main="Random Chart after Decomposition")
```

```{r}
#plotting each of the above
plot(decomMul$trend,xlab="Year",main="Trend Chart after Decomposition")
plot(decomMul$seasonal,xlab="Year",main="Seasonality Chart after Decomposition")
plot(decomMul$random,xlab="Year",main="Random Chart after Decomposition")
```

Subsetting the data

```{r}
DataAnalysis=window(data,start=1970)
start(DataAnalysis)
end(DataAnalysis)
frequency(DataAnalysis)
adf.test(DataAnalysis)
```


```{r}
plot(DataAnalysis,main="Plot of series from Year 1970",col='Blue')
abline(reg=lm(DataAnalysis~time(DataAnalysis)),col='red')

monthplot(DataAnalysis,main="Month plot of series from Year 1970 ",col='Blue')

boxplot(DataAnalysis~cycle(DataAnalysis),xlab = 'Month'
        , ylab = "Production",main="Boxplot for Australian Gas Production")

ggseasonplot(DataAnalysis,polar = TRUE)
gglagplot(DataAnalysis,main="Lagplot")

seasonplot(DataAnalysis,year.labels = TRUE,year.labels.left = TRUE,pch=19,col=1:40,
           main='Australian Monthly Gas Production-Seasonal Plot')


```

```{r}
decomAdd=decompose(DataAnalysis,type = "additive")
plot(DataAnalysis)
decomMul=decompose(DataAnalysis,type="multiplicative")
plot(decomAdd)
#plotting each of the above
plot(decomAdd$trend,xlab="Year",main="Trend Chart after Decomposition")
plot(decomAdd$seasonal,xlab="Year",main="Seasonality Chart after Decomposition")
plot(decomAdd$random,xlab="Year",main="Random Chart after Decomposition")

#plotting each of the above
plot(decomMul$trend,xlab="Year",main="Trend Chart after Decomposition")
plot(decomMul$seasonal,xlab="Year",main="Seasonality Chart after Decomposition")
plot(decomMul$random,xlab="Year",main="Random Chart after Decomposition")
plot(decomMul)
```


Null Hypothesis for ADF is Time series is Non-Stationery 
Alternate Hypothesis: Time series is sationary
Rejection of null hypothesis implies that series is stationary

```{r}
adf.test(DataAnalysis)
plot(DataAnalysis)
```

Removing seasonality

```{r}
decomp = stl(DataAnalysis, s.window = "periodic")
deseasonal_data=seasadj(decomp)
plot(decomp)
plot(DataAnalysis,col='blue',main="Seasonalized vs De-seasonalized plot")
lines(deseasonal_data,col='red' )
legend("bottomright", legend=c("Actual", "Deseasonalized"),col=c("blue", "red"), cex=0.8, lty= 1:1)

```

Removing Trend

```{r}
count_d1 = diff(deseasonal_data, differences = 1)
plot(count_d1, main="Differenced plot")
adf.test(count_d1, alternative = "stationary")

Acf(count_d1, main='ACF for Differenced Series',lag.max = 12)
Pacf(count_d1, main='PACF for Differenced Series',lag.max = 12)
```

Splitting the dataset into train and test

```{r}
dataTrain = window(deseasonal_data,start=1970,end=c(1993,12),frequency=12)
start(dataTrain)
end(dataTrain)
dataTest= window(deseasonal_data, start=1994,frequency=12)
start(dataTest)
end(dataTest)
```

Manually built ARIMA model
 
```{r}
plot(dataTrain)
arimaModel = arima(dataTrain, order=c(7,1,6))
arimaModel
tsdisplay(residuals(arimaModel),lag.max = 12, main='Model Residuals for Manual ARIMA')
```

Ljung box test
H0: Residuals are independent
Ha: Residuals are not independent

```{r}
library(stats)
Box.test(arimaModel$residuals)
```
 
Forecasting with manual arima for 20 periods
 
```{r}
fcast <- forecast(arimaModel, h=20,level = c(95))
plot(fcast)
```

Prediction with manual arima
 
```{r}
forecast::accuracy(fcast, dataTest,h=20)
fcast
#Actual vs Predicted on train dataset
pred <- predict(fcast, n.ahead = 20)
pred
plot(pred$x,col="blue",main="Actual vs Predicted on train dataset")
lines(pred$fitted,col='red' )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#Actual vs Predicted on test dataset
plot(DataAnalysis,col="blue",main="Actual vs Prediction on test Data" )
lines(fcast$mean,col="red" )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#comparing with test period
plot(pred$mean,col="blue",main="Forecasted test vs test Data" )
lines(dataTest,col="red" )

legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)
```

 
AUTO ARIMA Model

```{r pressure, echo=FALSE}
#partitioning seasonalized dataset
dataTrain1 = window(DataAnalysis,start=1970,end=c(1993,12),frequency=12)
start(dataTrain)
end(dataTrain)
dataTest1= window(DataAnalysis, start=1994,frequency=12)
```

```{r pressure, echo=FALSE}
#original trend and series model
fit1<-auto.arima(dataTrain1 , seasonal=TRUE)
fit1
tsdisplay(residuals(fit1), lag.max=12, main='Auto ARIMA Model Residuals')
#deseasonalized model
fit2=auto.arima(dataTrain,seasonal = TRUE)
fit2
tsdisplay(residuals(fit2), lag.max=12, main='Auto ARIMA Model Residuals')
```

Ljung box test
H0: Residuals are independent
Ha: Residuals are not independent

```{r}
Box.test(fit1$residuals)
Box.test(fit2$residuals)
```

forecasting with auto arima

```{r}
#original trend and series model
fcast1 <- forecast(fit1, h=20,level = c(95))
forecast::accuracy(fcast1,dataTest1)
plot(fcast1)
summary(fcast1)

#deseasonalized model
fcast2 <- forecast(fit2, h=20,level = c(95))
forecast::accuracy(fcast2,dataTest)
plot(fcast2)
summary(fcast2)


```


Plots of actual vs train and actual vs predicted 

```{r}
#original trend and series model

#Actual vs Predicted on train dataset
pred1 <- predict(fcast1, n.ahead = 20)
pred1
plot(pred1$x,col="blue",main="Actual vs Predicted on train dataset")
lines(pred1$fitted,col='red' )

legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#Actual vs Predicted on test dataset
plot(DataAnalysis,col="blue",main="Actual vs Prediction on test Data" )
lines(pred1$mean,col="red" )

legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#comparing with test period
plot(dataTest1,col="blue",main="Forecasted test vs test Data" )
lines(pred1$mean,col="red" )

legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)
```

```{r}
#deseasonalized series 
#Actual vs Predicted on train dataset
pred2 <- predict(fcast2, n.ahead = 20)
pred1
plot(pred2$x,col="blue",main="Actual vs Predicted on train dataset")
lines(pred2$fitted,col='red' )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#Actual vs Predicted on test dataset
plot(deseasonal_data,col="blue",main="Actual vs Prediction on test Data" )
lines(pred2$mean,col="red" )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#comparing with test period
plot(dataTest,col="blue",main="Forecasted test vs test Data" )
lines(pred2$mean,col="red" )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

```

Manual arima forecast for 20 periods
```{r}
f1=forecast(arimaModel,h=20)
forecast::accuracy(f1, dataTest)
summary(f1)
```
Auto airma forecast
```{r}
#original trend and series
f2=forecast(fit1,h=20)
forecast::accuracy(f2,dataTest1)
summary(f2)

```

```{r}
#deseasonalized
f3=forecast(fit2,h=20)
forecast::accuracy(f3,dataTest)
summary(f3)
```

Forecasting with 12 months

```{r}

fcast3 <- forecast(fit1, h=12,level = c(95))
forecast::accuracy(fcast3,dataTest)
plot(fcast3)
summary(fcast3)


predFinal <- predict(fcast3, n.ahead = 12)
pred1
plot(predFinal$x,col="blue",main="Actual vs Predicted on train dataset")
lines(predFinal$fitted,col='red' )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)

#Actual vs Predicted on test dataset
plot(DataAnalysis,col="blue",main="Actual vs Prediction on test Data" )
lines(predFinal$mean,col="red" )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)


#comparing with test period
plot(dataTest,col="blue",main="Forecasted test vs test Data" )
lines(predFinal$mean,col="red" )
legend("bottomright", legend=c("Actual", "Predicted"),col=c("blue", "red"), cex=0.8, lty= 1:1)
```

